---
title: "Project 1"
author: "Allison Shaffer"
format: html
---


```{r}
rm(list = ls())

library("here")
library("tidyverse")

if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}


if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}


chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

glimpse(chocolate)

colSums(is.na(chocolate))

```

Part 2
```{r}
#1
ggplot(chocolate, aes(x = rating)) +
  geom_histogram(bins = 25) +
  labs(
    title = "Distribution of Rating Scores",
    x = "Rating Score",
    y = "Count"
  )
```

Changing the number of bins changes the number of bars in the histogram (including the bars without any data). I chose 25 bins because it more clearly displays the breakdown of ratings between 3-4, where a majority of the observations fall. The 10 bin option was also appealing if we don't need to have as close of a look at the ratings. I think the option really depends on what you're trying to demonstrate with the graph. 

```{r}
#2
chocolate %>%
  count(country_of_bean_origin, sort = TRUE) %>%
  print(n = Inf)
```

```{r}
#3
chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  summarise(
    mean = mean(rating, na.rm = TRUE),
    sd = sd(rating, na.rm = TRUE),
    total = n()
  )
```

```{r}
#4
chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  group_by(company_location) %>%
  summarise(
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_rating))

```

Australia is the company location making the best chocolates with beans from Ecuador, with a mean rating of 3.81. 

```{r}
#5
chocolate %>%
  group_by(company_location) %>%
  summarise(
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_rating))

```

Chile, UAE, and Poland are the top 3 countries for bean origin with the highest mean rating, with rating of 3.75, 3.4, and 3.38 (respectively).


```{r}
#6
chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    total = n(),
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  filter(total >= 10) %>%
  arrange(desc(mean_rating)) %>%
  slice_head(n = 3)

```

Solomon Islands, Congo, and Cuba are the 3 countries with the highest mean rating, when only considering countries with 10 or greater reviews. 

```{r}
#7
chocolate_filtered <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 50) %>%
  ungroup()

chocolate_filtered <- chocolate_filtered %>%
  mutate(
    cocoa_percent_num = as.numeric(substr(cocoa_percent, 1, 2)),
    cocoa_group = case_when(
      cocoa_percent_num < 60 ~ "<60%",
      cocoa_percent_num >= 60 & cocoa_percent_num < 70 ~ "60-<70%",
      cocoa_percent_num >= 70 & cocoa_percent_num < 90 ~ "70-<90%",
      cocoa_percent_num >= 90 ~ ">=90%"
    )
  )

chocolate_filtered <- chocolate_filtered %>%
  mutate(
    cocoa_group = fct_relevel(
      cocoa_group,
      "<60%",
      "60-<70%",
      "70-<90%",
      ">=90%"
    )
  )

ggplot(chocolate_filtered, aes(x = cocoa_group, y = rating)) +
  geom_boxplot() +
  facet_wrap(~ country_of_bean_origin) +
  labs(
    title = "Chocolate Ratings by Cocoa Percentage and Country",
    x = "Chocolate Percentage",
    y = "Rating"
  )

```

Based on a visual review of the plots, the chocolates with cocoa percentages between 60 & 70 and 70 & 90 seem to have better ratings, but there are differences between country of origin 


Part 2
```{r}
library(tidyverse)
library(gapminder)

#1
chocolate <- chocolate %>%
  left_join(
    gapminder %>%
      select(country, continent) %>%
      distinct(),
    by = c("country_of_bean_origin" = "country")
  )

sum(is.na(chocolate$continent))

chocolate %>%
  filter(is.na(continent)) %>%
  count(country_of_bean_origin, sort = TRUE) %>%
  View()

chocolate <- chocolate %>%
  mutate(
    continent = case_when(
      country_of_bean_origin %in% c("Congo", "Ivory Coast", "Sao Tome", "Sao Tome & Principe", "DR Congo", "Principe") ~ "Africa",
      country_of_bean_origin %in% c("Belize", "Trinidad", "U.S.A.", "Grenada", "St. Lucia", "St.Vincent-Grenadines", "Martinique", "Suriname", "Tobago") ~ "Americas",
      country_of_bean_origin %in% c("Burma", "Sulawesi", "Sumatra") ~ "Asia",
      country_of_bean_origin %in% c("Fiji", "Papua New Guinea", "Vanuatu", "Solomon Islands", "Samoa") ~ "Oceania",
      TRUE ~ continent
    )
  )


sum(is.na(chocolate$continent))

chocolate %>%
  filter(is.na(continent)) %>%
  count(country_of_bean_origin, sort = TRUE) %>%
  View()

#2 
chocolate <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10) %>%
  ungroup()

#3
chocolate <- chocolate %>%
  filter(country_of_bean_origin != "Blend")


#4
ggplot(chocolate, aes(x = continent, y = rating)) +
  geom_violin() +
  labs(
    title = "Chocolate Ratings by Continent",
    x = "Continent",
    y = "Rating"
  )
```


Part 3 
```{r}
#1
chocolate <- chocolate %>%
  mutate(
    beans = if_else(str_detect(ingredients, "B"), 1, 0),
    sugar = if_else(str_detect(ingredients, "S") & !str_detect(ingredients, "S\\*|Sa"), 1, 0),
    cocoa_butter = if_else(str_detect(ingredients, "C"), 1, 0),
    vanilla = if_else(str_detect(ingredients, "V"), 1, 0),
    lecithin = if_else(str_detect(ingredients, "L"), 1, 0),
    salt = if_else(str_detect(ingredients, "Sa"), 1, 0)
  )

#2
chocolate <- chocolate %>%
  mutate(
    char_cocoa = if_else(str_detect(most_memorable_characteristics, "cocoa"), 1, 0),
    char_sweet = if_else(str_detect(most_memorable_characteristics, "sweet"), 1, 0),
    char_nutty = if_else(str_detect(most_memorable_characteristics, "nutty"), 1, 0),
    char_creamy = if_else(str_detect(most_memorable_characteristics, "creamy"), 1, 0),
    char_roasty = if_else(str_detect(most_memorable_characteristics, "roasty"), 1, 0),
    char_earthy = if_else(str_detect(most_memorable_characteristics, "earthy"), 1, 0)
  )

 #3 

 chocolate_yearly <- chocolate %>%
  group_by(review_date) %>%
  summarise(
    beans = mean(beans, na.rm = TRUE),
    sugar = mean(sugar, na.rm = TRUE),
    cocoa_butter = mean(cocoa_butter, na.rm = TRUE),
    vanilla = mean(vanilla, na.rm = TRUE),
    lecithin = mean(lecithin, na.rm = TRUE),
    salt = mean(salt, na.rm = TRUE),
    char_cocoa = mean(char_cocoa, na.rm = TRUE),
    char_sweet = mean(char_sweet, na.rm = TRUE),
    char_nutty = mean(char_nutty, na.rm = TRUE),
    char_creamy = mean(char_creamy, na.rm = TRUE),
    char_roasty = mean(char_roasty, na.rm = TRUE),
    char_earthy = mean(char_earthy, na.rm = TRUE)
  )
chocolate_yearly

#4
chocolate_yearly_long <- chocolate_yearly %>%
  pivot_longer(
    cols = -review_date,
    names_to = "feature",
    values_to = "mean_score"
  )
chocolate_yearly_long

```


```{r}
#Part 4
chocolate_yearly_long <- chocolate_yearly_long %>%
  mutate(
    feature = factor(
      feature,
      levels = c(
        "beans",
        "sugar",
        "char_cocoa",
        "char_sweet",
        "cocoa_butter",
        "lecithin",
        "char_nutty",
        "char_creamy",
        "vanilla",
        "salt",
        "char_roasty",
        "char_earthy"
      )
    )
  )

ggplot(chocolate_yearly_long, 
       aes(x = review_date, y = mean_score, color = feature)) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  facet_wrap(~ feature) +
  scale_color_manual(
    values = c(
      "beans" = "blue",
      "sugar" = "blue",
      "cocoa_butter" = "blue",
      "vanilla" = "blue",
      "lecithin" = "blue",
      "salt" = "blue",
      "char_cocoa" = "red",
      "char_sweet" = "red",
      "char_nutty" = "red",
      "char_creamy" = "red",
      "char_roasty" = "red",
      "char_earthy" = "red"
    )
  ) +
  labs(
    title = "Chocolate Ingredients and Characteristics Over Time",
    subtitle = "The majority of ingredients and characteristics were relatively stable over time. Lecithin, vanilla, and cocoa butter have notable\ndownward trends throughout the years.",
    x = "Review Year",
    y = "Mean Score",
    caption = "Allison Shaffer"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 8),
    strip.text = element_text(face = "bold"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

```  


```{r}

#Part 5 

ggplot(chocolate, aes(x = cocoa_percent, y = rating, color = company_location)) +
  geom_point(size = 5, alpha = 0.8) +
  geom_smooth(method = "lm", linewidth = 3, se = TRUE) +
  facet_wrap(~ review_date) +
  scale_color_manual(values = rainbow(length(unique(chocolate$company_location)))) +
  labs(
    title = "THE ABSOLUTELY AMAZING SUPER IMPORTANT CHOCOLATE RATING PLOT!!!",
    subtitle = "This visualization contains WAY too much information and is definitely not confusing at all!!!",
    x = "COCOA PERCENTAGE OF CHOCOLATE",
    y = "RATING SCORE",
    caption = "Created by Allison Shaffer — The World's Most Questionable Data Visualization"
  ) +
  theme(
    panel.background = element_rect(fill = "yellow"),
    plot.background = element_rect(fill = "pink"),
    panel.grid.major = element_line(color = "red", linewidth = 2),
    panel.grid.minor = element_line(color = "blue", linewidth = 1),
    axis.text.x = element_text(angle = 90, size = 16, color = "purple"),
    axis.text.y = element_text(size = 16, color = "orange"),
    axis.title = element_text(size = 20, color = "green"),
    plot.title = element_text(size = 24, color = "red", face = "bold"),
    plot.subtitle = element_text(size = 16, color = "blue"),
    strip.background = element_rect(fill = "limegreen"),
    strip.text = element_text(size = 16, color = "white"),
    legend.background = element_rect(fill = "cyan"),
    legend.text = element_text(size = 14, color = "purple")
  )
```


```{r}

#Part 6 
chocolate %>%
  ggplot(aes(
    x = as.factor(review_date),
    y = rating,
  )) +
  geom_violin(alpha = 0.7, fill = "steelblue") +
  labs(
    title = "Chocolate Ratings by Review Year",
    x = "Review Year",
    y = "Rating",
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

#1 I removed the varying color scale, the different colors didn't add any additional information and were distracting
#2 I made the color or the plots steel blue and slightly transparent for visual appeal. I
#3 I added a title, which helps a viewer more easily understand what this graph is showing.
#4 I added x and y axis labels that had more meaning. 
#5 I tilted and aligned the year labels along the x axis so it doesn't appear as crowded
#6 I made the theme minimal to remove the appearance (removed gray background and replaced with white) 
#7 I also removed the legend as they're all one color now. 




#of note, i ran my analysis and my graphs on my cleaned data (from part 2), while it seems perhaps yours is the raw data, having not removed the 'blends'. 

```