---
title: "Project1"
format: html
---


```{r}
rm(list = ls())

library("here")
library("tidyverse")

if (!dir.exists(here("data"))) {
    dir.create(here("data"))
}


if (!file.exists(here("data", "chocolate.RDS"))) {
    url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2022/2022-01-18/chocolate.csv"
    chocolate <- readr::read_csv(url_csv)

    # save the file to RDS objects
    saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}


chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)


glimpse(chocolate)


colSums(is.na(chocolate))

```


```{r}
#Part 2
#1
ggplot(chocolate, aes(x = rating)) +
  geom_histogram(bins = 25) +
  labs(
    title = "Distribution of Rating Scores",
    x = "Rating Score",
    y = "Count"
  )

  #Changing the number of bins changes the number of bars in the histogram (including the bars without any data). I chose 25 bins because it more clearly displays the breakdown of ratings between 3-4, where a majority of the observations fall. The 10 bin option was also appealing if we don't need to have as close of a look at the ratings. I think the option really depends on what you're trying to demonstrate with the graph. 

```

```{r}
#2
chocolate %>%
  count(country_of_bean_origin, sort = TRUE)
```

```{r}
#3
chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  summarise(
    mean = mean(rating, na.rm = TRUE),
    sd = sd(rating, na.rm = TRUE),
    total = n()
  )
```

```{r}
#4
chocolate %>%
  filter(country_of_bean_origin == "Ecuador") %>%
  group_by(company_location) %>%
  summarise(
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_rating))

#Australia
```


```{r}
#5
chocolate %>%
  group_by(company_location) %>%
  summarise(
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  arrange(desc(mean_rating))

#Chile 
```

```{r}

#6
chocolate %>%
  group_by(country_of_bean_origin) %>%
  summarise(
    total = n(),
    mean_rating = mean(rating, na.rm = TRUE)
  ) %>%
  filter(total >= 10) %>%
  arrange(desc(mean_rating)) %>%
  slice_head(n = 3)

#Solomon Islands, Congo, Cuba 

```

```{r}
#7
chocolate_filtered <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 50) %>%
  ungroup()

chocolate_filtered <- chocolate_filtered %>%
  mutate(
    cocoa_percent_num = as.numeric(substr(cocoa_percent, 1, 2)),
    cocoa_group = case_when(
      cocoa_percent_num < 60 ~ "<60%",
      cocoa_percent_num >= 60 & cocoa_percent_num < 70 ~ "60-<70%",
      cocoa_percent_num >= 70 & cocoa_percent_num < 90 ~ "70-<90%",
      cocoa_percent_num >= 90 ~ ">=90%"
    )
  )

chocolate_filtered <- chocolate_filtered %>%
  mutate(
    cocoa_group = fct_relevel(
      cocoa_group,
      "<60%",
      "60-<70%",
      "70-<90%",
      ">=90%"
    )
  )

ggplot(chocolate_filtered, aes(x = cocoa_group, y = rating)) +
  geom_boxplot() +
  facet_wrap(~ country_of_bean_origin) +
  labs(
    title = "Chocolate Ratings by Cocoa Percentage and Country",
    x = "Chocolate Percentage",
    y = "Rating"
  )

#Based on a visual review of the plots, the chocolates with cocoa percentages between 60 & 70 and 70 & 90 seem to have better ratings, but there are differences between country of origin  

```

```{r}
#Part 2
library(tidyverse)
library(gapminder)
gapminder


#1
chocolate <- chocolate %>%
  left_join(
    gapminder %>%
      select(country, continent) %>%
      distinct(),
    by = c("country_of_bean_origin" = "country")
  )

sum(is.na(chocolate$continent))

chocolate %>%
  filter(is.na(continent)) %>%
  count(country_of_bean_origin, sort = TRUE) %>%
  View()

chocolate <- chocolate %>%
  mutate(
    continent = case_when(
      country_of_bean_origin %in% c("Congo", "Ivory Coast", "Sao Tome", "Sao Tome & Principe", "DR Congo", "Principe") ~ "Africa",
      country_of_bean_origin %in% c("Belize", "Trinidad", "U.S.A.", "Grenada", "St. Lucia", "St.Vincent-Grenadines", "Martinique", "Suriname", "Tobago") ~ "Americas",
      country_of_bean_origin %in% c("Burma", "Sulawesi", "Sumatra") ~ "Asia",
      country_of_bean_origin %in% c("Fiji", "Papua New Guinea", "Vanuatu", "Solomon Islands", "Samoa") ~ "Oceania",
      TRUE ~ continent
    )
  )


sum(is.na(chocolate$continent))

chocolate %>%
  filter(is.na(continent)) %>%
  count(country_of_bean_origin, sort = TRUE) %>%
  View()

#2 
chocolate <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10) %>%
  ungroup()

#3
chocolate <- chocolate %>%
  filter(country_of_bean_origin != "Blend")


#4
chocolate <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 10) %>%
  ungroup()

```



```{r}
#Part 3 

#1
chocolate <- chocolate %>%
  mutate(
    beans = if_else(str_detect(ingredients, "B"), 1, 0),
    sugar = if_else(str_detect(ingredients, "S") & !str_detect(ingredients, "S\\*|Sa"), 1, 0),
    cocoa_butter = if_else(str_detect(ingredients, "C"), 1, 0),
    vanilla = if_else(str_detect(ingredients, "V"), 1, 0),
    lecithin = if_else(str_detect(ingredients, "L"), 1, 0),
    salt = if_else(str_detect(ingredients, "Sa"), 1, 0)
  )

#2
chocolate <- chocolate %>%
  mutate(
    char_cocoa = if_else(str_detect(most_memorable_characteristics, "cocoa"), 1, 0),
    char_sweet = if_else(str_detect(most_memorable_characteristics, "sweet"), 1, 0),
    char_nutty = if_else(str_detect(most_memorable_characteristics, "nutty"), 1, 0),
    char_creamy = if_else(str_detect(most_memorable_characteristics, "creamy"), 1, 0),
    char_roasty = if_else(str_detect(most_memorable_characteristics, "roasty"), 1, 0),
    char_earthy = if_else(str_detect(most_memorable_characteristics, "earthy"), 1, 0)
  )

 #3 

 chocolate_yearly <- chocolate %>%
  group_by(review_date) %>%
  summarise(
    beans = mean(beans, na.rm = TRUE),
    sugar = mean(sugar, na.rm = TRUE),
    cocoa_butter = mean(cocoa_butter, na.rm = TRUE),
    vanilla = mean(vanilla, na.rm = TRUE),
    lecithin = mean(lecithin, na.rm = TRUE),
    salt = mean(salt, na.rm = TRUE),
    char_cocoa = mean(char_cocoa, na.rm = TRUE),
    char_sweet = mean(char_sweet, na.rm = TRUE),
    char_nutty = mean(char_nutty, na.rm = TRUE),
    char_creamy = mean(char_creamy, na.rm = TRUE),
    char_roasty = mean(char_roasty, na.rm = TRUE),
    char_earthy = mean(char_earthy, na.rm = TRUE)
  )


chocolate_yearly_long <- chocolate_yearly %>%
  pivot_longer(
    cols = -review_date,
    names_to = "feature",
    values_to = "mean_score"
  )
